---
title: 'Data Wrangling Ex 1: Basic Data Manipulation'
author: "Anthony Stachowitz"
date: "April 20, 2018"
output: html_document
---

```{r, message=FALSE}
library("plyr")
library("dplyr")
library("tidyr")
library("readxl")
library("knitr")
```

# Exercise

Using R, clean the data set ***refine*** to make it easier to visualize and analyze.

## Steps

### 1) Load the data in RStudio:

#### Save the data set as a CSV file called refine_original.csv and load it in RStudio into a data frame.

```{r}
refine_original <- read_excel("C:/Users/xena0/Downloads/refine.xlsx")
View(refine_original)
```

### 2) Clean up brand names

#### Clean up the 'company' column so all of the misspellings of the brand names are standardized. For example, you can transform the values in the column to be: philips, akzo, van houten and unilever (all lowercase).

```{r}
refine_clean <- refine_original %>% 
  mutate(company = ifelse(grepl("^phil|^fil|^phl", company, ignore.case = TRUE), "philips", company)) %>% 
  mutate(company = ifelse(grepl("ak", company, ignore.case = TRUE), "akzo", company)) %>% 
  mutate(company = ifelse(grepl("van", company, ignore.case = TRUE), "van houten", company)) %>% 
  mutate(company = ifelse(grepl("uni", company, ignore.case = TRUE), "unilever", company))
```

### 3) Separate product code and number

#### Separate the product code and product number into separate columns i.e. add two new columns called product_code and product_number, containing the product code and number respectively

```{r}
refine_clean <- refine_clean %>%
  separate(col = "Product code / number",
          into = c("product_code", "product_number"),
          sep = "-")
```

### 4) Add product categories

####   You learn that the product codes actually represent the following product categories:

####    p = Smartphone

####    v = TV

####    x = Laptop

####    q = Tablet

####   In order to make the data more readable, add a column with the product category for each record.

```{r}
refine_clean <- mutate(refine_clean, product_categories = 0)

refine_clean <- refine_clean %>% 
  mutate(product_categories = ifelse(grepl("p", product_code), "Smartphone", product_categories)) %>% 
  mutate(product_categories = ifelse(grepl("v", product_code), "TV", product_categories)) %>% 
  mutate(product_categories = ifelse(grepl("x", product_code), "Laptop", product_categories)) %>% 
  mutate(product_categories = ifelse(grepl("q", product_code), "Tablet", product_categories))
```

### 5) Add full address for geocoding

#### You'd like to view the customer information on a map. In order to do that, the addresses need to be in a form that can be easily geocoded. Create a new column full_address that concatenates the three address fields (address, city, country), separated by commas.

```{r}
refine_clean <- refine_clean %>% 
  unite(full_address, address:country, sep = ", ", remove = FALSE)
```

### 6) Create dummy variables for company and product category

#### Both the company name and product category are categorical variables i.e. they take only a fixed set of values. In order to use them in further analysis you need to create dummy variables. Create dummy binary variables for each of them with the prefix company_ and product_ i.e.,

#### Add four binary (1 or 0) columns for company: company_philips, company_akzo, company_van_houten and company_unilever.

#### Add four binary (1 or 0) columns for product category: product_smartphone, product_tv, product_laptop and product_tablet.

```{r}
refine_clean <- refine_clean %>% 
  mutate(company_philips = ifelse(grepl("philips", company), 1, 0)) %>% 
  mutate(company_akzo = ifelse(grepl("akzo", company),  1, 0)) %>% 
  mutate(company_van_houten = ifelse(grepl("van houten", company),  1, 0)) %>% 
  mutate(company_unilever = ifelse(grepl("unilever", company),  1, 0)) %>% 
  mutate(product_smartphone = ifelse(grepl("Smartphone", product_categories),  1, 0)) %>% 
  mutate(product_tv = ifelse(grepl("TV", product_categories),  1, 0)) %>% 
  mutate(product_laptop = ifelse(grepl("Laptop", product_categories),  1, 0)) %>% 
  mutate(product_tablet = ifelse(grepl("Tablet", product_categories),  1, 0))
```

### 6) Submit the project on Github

#### Include your code, the original data as a CSV file refine_original.csv, and the cleaned up data as a CSV file called refine_clean.csv.

```{r}
kable(refine_clean[1:5, ], caption = "refine clean")
```
